Estimation of Intracortical Connectivity in a Dynamical Neural Field Model

### 2010-04-13 ltx/BrainIDE.tex
added to freq analysis in paper

There are two options available to alleviate problems with aliasing. The first is to increase the spatial sampling frequency (using a higher resolution of sensors), and the second is to use wider sensors, which act like an anti-aliasing filter. The first option is more desirable, since more information is retained. In addition, if sensors are modelled as having a Gaussian pickup range than the slope of the attenuation will be gentle allowing for some low amplitude high frequency components to still be present and mild aliasing will still occur.  Mild aliasing results in distortions in the high frequency components of the field dynamics. The basis function decomposition the high frequency components are usually lost.

Similar considerations need to be made regarding the representation of the dynamic field $w_t(\mathbf{r})$ by its basis functions decompositions. The minimum distance between adjacent basis needs to again satisfy Shannon's sampling theorem. Thus the minimum distance between basis must again satisfy $\frac{1}{2\nu_c}$. Note that, since the dynamic field is frequently spatially oversampled by using a larger number of sensor then imposed by frequency considerations, the distance between basis functions is usually larger then the distance between sensors. Note also that since a limited spatial region and since basis functions placed on a regular square grid are being considered, knowledge of the distance between basis functions directly implies the total number of basis functions required.

For the Gaussian basis functions $\phi(\mathbf{r})$ being considered, the basis width parameter $\sigma^2$ can also be inferred from spectral considerations \cite{Sanner1992,Scerri2009}. The Fourier transform of a Gaussian is another Gaussian with width $\sigma^2_{\nu}$ given by $\sigma^2_{\nu} = \frac{1}{\pi^2\sigma^2}$. Thus to ensure that the basis functions can represent fields with frequency content up to $\boldsymbol{\nu}_c$, $\sigma_{\nu}$ should be set with the 3dB frequency for the basis functions equal to $\boldsymbol{\nu}_c$, thus obtained $\sigma^2_{\nu}$ and therefore $\sigma^2$.

>>>>

There are two options available to alleviate problems with aliasing. The first is to increase the spatial sampling frequency (using a higher resolution of sensors), and the second is to use wider sensors, which act like an anti-aliasing filters. The first option is more desirable, since more information is retained. If sensors are modelled as having a Gaussian pickup range than the slope of the attenuation will be gentle allowing for some low amplitude high frequency components to still be present and mild aliasing will still occur. In practice, the bandwidth of the cortex can not be estimated by traditional electrophysiological measurements due to spatial limitations. It may be estimated using other modalities with higher spatial resolution such as, fMRI, NIRS or other optical imaging techniques. Errors associated with mild aliasing result in distortions in the high frequency components of the field dynamics. However, after applying the basis function decomposition, high frequency component contribute negligibly to the inverse problem and are therefore not a major problem~\cite{Sanner1992}. This is further illustrated our results.

Similar considerations need to be made regarding the representation of the dynamic field $w_t(\mathbf{r})$ by its basis functions decompositions. The minimum distance between adjacent basis functions needs to again satisfy Shannon's sampling theorem. Thus the minimum distance between basis must again satisfy $\frac{1}{2\nu_c}$. For the Gaussian basis functions $\phi(\mathbf{r})$ being considered, the basis width parameter $\sigma^2$ can also be inferred from spectral considerations \cite{Sanner1992,Scerri2009}. The Fourier transform of a Gaussian is another Gaussian with width $\sigma^2_{\nu}$ given by $\sigma^2_{\nu} = \frac{1}{\pi^2\sigma^2}$. Thus to ensure that the basis functions can represent fields with frequency content up to $\boldsymbol{\nu}_c$, $\sigma_{\nu}$ should be set with the 3dB frequency for the basis functions equal to $\boldsymbol{\nu}_c$, thus obtained $\sigma^2_{\nu}$ and therefore $\sigma^2$.

The dynamic field must be spatially oversampled by using a larger number of sensor then imposed by frequency considerations, to use frequency analysis to determine the minimum distance and width of the basis functions. Following this, the distance between basis functions is usually larger then the distance between sensors. \todo{However this is not the case with this work.} For a spatially homogeneous isotropic field the basis functions can be placed on a regular square grid. Following this, the knowledge of the distance between basis functions directly implies the total number of basis functions required.

Due to the high dimensionality of the brain and the current electrode systems, we can not expect to have more sensors than basis functions.
